---
title: "chapter-5-data-transformation"
output: html_document
---

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
install.packages("tidyverse", repos="http://cran.us.r-project.org")
library(tidyverse)

install.packages("nycflights13", repos="http://cran.us.r-project.org")
library(nycflights13)

library(ggplot2)
```

# Chapter 5: Data transformation

```{r}
flights
```

# 5.2: Filter rows with `filter()`

```{r}
filter(flights, month == 1, day == 1)
```

```{r}
filter(flights, month == 11 | month == 12)
```

```{r}
filter(flights, month %in% c(11, 12))
```

## Exercises

1.1 

```{r}
filter(flights, arr_delay >= 120)
```

1.2

```{r}
filter(flights, dest %in% c("IAH", "HOU"))
```

1.3

```{r}
airlines
```


```{r}
filter(flights, carrier %in% c("UA", "AA", "DL"))
```

1.4

```{r}
filter(flights, month %in% c(7, 8, 9))
```

1.5

```{r}
filter(flights, dep_delay <= 0 & arr_delay > 120)
```

1.6 

```{r}
filter(flights, dep_delay >= 60 & arr_delay < 30)
```

1.7

```{r}
filter(flights, dep_time >=0 & dep_time <= 600)
```

2.

```{r}
filter(flights, between(dep_time, 0, 600))
```

3.

```{r}
filter(flights, is.na(dep_time))
```

4.

`NA` and logical operators:

```{r}
NA | TRUE
```

```{r}
NA | FALSE
```

```{r}
NA & FALSE
```

```{r}
NA & TRUE
```

```{r}
x <- c(TRUE, FALSE, NA)
```
```{r}
names(x) <- as.character(x)
```

`AND` table:
```{r}
outer(x, x, "&")
```

`OR` table:

```{r}
outer(x, x, "|")
```


`NA` and computational operators:

```{r}
NA * 0
```


```{r}
NA ^ 0
```

# 5.3: Arrange rows with `arrange()`

```{r}
arrange(flights, year, month, day)
```

```{r}
arrange(flights, desc(arr_delay))
```

Missing values:

```{r}
df <- tibble(x = c(5, 2, NA))
```
```{r}
arrange(df, x)
```

```{r}
arrange(df, desc(x))
```

## Exercises

1.

```{r}
df %>% arrange(!is.na(x), x)
```

```{r}
df_with_logical_values = tibble(x = c(TRUE, FALSE, NA))
```
```{r}
df_with_logical_values %>% arrange(x)
```

```{r}
df_with_logical_values %>% arrange(desc(x))
```

### 2.1 Most delayed flights

```{r}
flights %>% arrange(desc(dep_delay))
```

### 2.2 Flights that left the earliest

```{r}
flights %>% arrange(dep_delay)
```

### 3. Fastest flights

```{r}
flights %>% arrange(air_time)
```

### 4.1 Flights travelling the longest

```{r}
flights %>% arrange(desc(distance))
```


### 4.2 Flights traevlling the shortest

```{r}
flights %>% arrange(distance)
```



# 5.4 Select columns with `select()`

```{r}
flights %>% select(year, month, day)
```

```{r}
flights %>% select(year:day)
```

```{r}
flights %>% select(-(year:day))
```
```{r}
flights %>% select(tail_num = tailnum)
```

```{r}
flights %>% rename(tail_num = tailnum)
```

```{r}
flights %>% select(time_hour, air_time, everything())
```

## Exercises

### 1.

```{r}
flights %>% select(dep_time, dep_delay, arr_time, arr_delay)
```

```{r}
flights %>% select(dep_time:arr_delay, -sched_dep_time, -sched_arr_time)
```

```{r}
flights %>% select(c('dep_time', 'dep_delay', 'arr_time', 'arr_delay'))
```

```{r}
flights %>% select(starts_with('dep_'), starts_with('arr_'))
```

```{r}
flights %>% select(contains("_time"), contains("_delay"))
```

### 2. Including a variable name multiple times

```{r}
flights %>% select(arr_time, arr_time)
```

### 3. `one_of()` function

```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
```

```{r}
flights %>% select(one_of(vars))
```

### 4

```{r}
select(flights, contains("TIME"))
```

```{r}
flights %>% select(contains("TIME", ignore.case = FALSE))
```

# 5.5 Add new variables with `mutate()`

```{r}
flights_small <- select(flights, year:day, ends_with("delay"), distance, air_time)
```

```{r}
flights_small %>% mutate(
  gain = arr_delay - dep_delay,
  speed = distance / air_time * 60
)
```
