---
title: "r-for-data-science"
output: html_document
---

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
install.packages("tidyverse", repos="http://cran.us.r-project.org")
library(tidyverse)

library(ggplot2)
```


# Chapter 1

## First Steps

### Engine Displacement (in Litres) by Highway Mileage (per Gallon)

```{r}
ggplot(data = ggplot2::mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))
```

### 1

> Run ggplot(data = mpg). What do you see?

Blank canvas.

```{r}
ggplot(data = ggplot2::mpg)
```

### 2

> How many rows are in mtcars? How many columns?

```{r}
ggplot2::mpg
```

### 3

> What does the drv variable describe? Read the help for ?mpg to find out.

```{r}
?mpg
```

### 4

> Make a scatterplot of hwy versus cyl.

```{r}
ggplot(data = ggplot2::mpg) +
  geom_point(mapping = aes(x = hwy, y = cyl))
```

```{r}
ggplot(data = ggplot2::mpg) +
  geom_jitter(mapping = aes(x = hwy, y = cyl))
```

```{r}
ggplot(data = ggplot2::mpg) +
  geom_count(mapping = aes(x = hwy, y = cyl))
```



```{r}
ggplot(data = ggplot2::mpg) +
  geom_bin2d(mapping = aes(x = hwy, y = cyl))
```

```{r}
ggplot(data = ggplot2::mpg) +
  geom_point(mapping = aes(x = hwy, y = cyl), alpha = 0.05)
```

### 5

> What happens if you make a scatterplot of class versus drv? Why is the plot not useful?

`class` and `drv` are both categorical variables.

```{r}
ggplot(data = ggplot2::mpg) +
  geom_point(mapping = aes(x = class, y = drv))
```


## Aesthetic Mappings

### Engine Displacement (in Litres) by Highway Mileage (per Gallon), by Class (Type of Car)

```{r}
ggplot(data = ggplot2::mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
```

### 1

```{r}
ggplot2::mpg
```

### 2

> Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset.) How can you see this information when you run mpg?

```{r}
str(ggplot2::mpg)
```

### 3

> Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical versus continuous variables?

```{r}
ggplot(data = ggplot2::mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = year))
```

### 4

> What happens if you map the same variable to multiple aesthetics?


```{r}
ggplot(data = ggplot2::mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = hwy))
```

### 5

> What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point.)

```{r}
ggplot(data = ggplot2::mpg) +
  geom_point(mapping = aes(x = displ, y = hwy), shape = 21, colour = "black", fill = "white", size = 5, stroke = 5)
```

### 6

> What happens if you map an aesthetic to something other than a variable name, like aes(color = displ < 5)?

```{r}
ggplot(data = ggplot2::mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = displ < 5))
```

## Common Problems

```{r}
ggplot(data = mpg)  
  + geom_point(mapping = aes(x = displ, y = hwy))
```