---
title: "r-for-data-science"
output: html_document
---

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
install.packages("tidyverse", repos="http://cran.us.r-project.org")
library(tidyverse)

library(ggplot2)
```


# Chapter 1

## First Steps

### Engine Displacement (in Litres) by Highway Mileage (per Gallon)

```{r}
ggplot(data = ggplot2::mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))
```

### Exercises

#### 1

> Run ggplot(data = mpg). What do you see?

Blank canvas.

```{r}
ggplot(data = ggplot2::mpg)
```

#### 2

> How many rows are in mtcars? How many columns?

```{r}
ggplot2::mpg
```

#### 3

> What does the drv variable describe? Read the help for ?mpg to find out.

```{r}
?mpg
```

#### 4

> Make a scatterplot of hwy versus cyl.

```{r}
ggplot(data = ggplot2::mpg) +
  geom_point(mapping = aes(x = hwy, y = cyl))
```

```{r}
ggplot(data = ggplot2::mpg) +
  geom_jitter(mapping = aes(x = hwy, y = cyl))
```

```{r}
ggplot(data = ggplot2::mpg) +
  geom_count(mapping = aes(x = hwy, y = cyl))
```



```{r}
ggplot(data = ggplot2::mpg) +
  geom_bin2d(mapping = aes(x = hwy, y = cyl))
```

```{r}
ggplot(data = ggplot2::mpg) +
  geom_point(mapping = aes(x = hwy, y = cyl), alpha = 0.05)
```

#### 5

> What happens if you make a scatterplot of class versus drv? Why is the plot not useful?

`class` and `drv` are both categorical variables.

```{r}
ggplot(data = ggplot2::mpg) +
  geom_point(mapping = aes(x = class, y = drv))
```


## Aesthetic Mappings

### Engine Displacement (in Litres) by Highway Mileage (per Gallon), by Class (Type of Car)

```{r}
ggplot(data = ggplot2::mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
```

### Exercises

#### 1

```{r}
ggplot2::mpg
```

#### 2

> Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset.) How can you see this information when you run mpg?

```{r}
str(ggplot2::mpg)
```

#### 3

> Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical versus continuous variables?

```{r}
ggplot(data = ggplot2::mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = year))
```

#### 4

> What happens if you map the same variable to multiple aesthetics?


```{r}
ggplot(data = ggplot2::mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = hwy))
```

#### 5

> What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point.)

```{r}
ggplot(data = ggplot2::mpg) +
  geom_point(mapping = aes(x = displ, y = hwy), shape = 21, colour = "black", fill = "white", size = 5, stroke = 5)
```

#### 6

> What happens if you map an aesthetic to something other than a variable name, like aes(color = displ < 5)?

```{r}
ggplot(data = ggplot2::mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = displ < 5))
```

## Common Problems

```{r}
#ggplot(data = mpg)  
#  + geom_point(mapping = aes(x = displ, y = hwy))
```

## Facets

### Highway Mileage (per Gallon) by Engine Displacement (in Litres), By Car Type

```{r}
ggplot(data = ggplot2::mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

### Highway Mileage (per Gallon) by Engine Displacement (in Litres), By Drive Type, By Number of Cylinders

```{r}
ggplot(data = ggplot2::mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ cyl)
```


### Highway Mileage (per Gallon) by Engine Displacement (in Litres), By Number of Cylinders

```{r}
ggplot(data = ggplot2::mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```

### Exercises

#### 1

> What happens if you facet on a continuous variable?

```{r}
ggplot(data = ggplot2::mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cty)
```

#### 2

> What do the empty cells in a plot with facet_grid(drv ~ cyl) mean? How do they relate to this plot?

```{r}
ggplot(data = ggplot2::mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ cyl)
```

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = drv, y = cyl))
```

#### 3

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)
```

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```

#### 4

> Take the first faceted plot in this section:

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~ class, nrow = 2)
```

> What are the advantages to using faceting instead of the color aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?

```{r}
ggplot(data = ggplot2::mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(~ class)
```


```{r}
ggplot(data = ggplot2::mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
```

#### 5

```{r}
?facet_wrap
```

## Geometric Objects

### Highway Mileage (per Gallon) by Engine Displacement (in Litres)

```{r}
ggplot(data = ggplot2::mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))
```

```{r}
ggplot(data = ggplot2::mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
```

### Highway Mileage (per Gallon) by Engine Displacement (in Litres), By Drive Type

```{r}
ggplot(data = ggplot2::mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))
```

```{r}
ggplot(data = ggplot2::mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv, color = drv)) +
  geom_point(mapping = aes(x = displ, y = hwy, color = drv))
```


```{r}
ggplot(data = ggplot2::mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy, group = drv))

```

```{r}
ggplot(data = mpg) +
  geom_smooth(
    mapping = aes(x = displ, y = hwy, color = drv),
    show.legend = FALSE
  )
```

```{r}
ggplot(data = ggplot2::mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
```

```{r}
ggplot(data = ggplot2::mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point() +
  geom_smooth()
```

```{r}
ggplot(data = ggplot2::mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(color = class)) + 
  geom_smooth()
```

```{r}
ggplot(data = ggplot2::mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(color = class)) + 
  geom_smooth(
    data = filter(mpg, class == "subcompact"),
    se = FALSE,
    mapping = aes(color = class)
  )
```

### Exercises

#### 1



#### 2

```{r}
ggplot(
  data = mpg,
  mapping = aes(x = displ, y = hwy, color = drv)
) +
  geom_point() +
  geom_smooth(se = FALSE)
```

#### 4

```{r}
ggplot(
  data = ggplot2::mpg,
  mapping = aes(x = displ, y = hwy, color = drv)
) +
  geom_point() +
  geom_smooth(se = TRUE)
```

## Statistical Transformations

```{r}
ggplot(data = ggplot2::diamonds) +
  geom_bar(mapping = aes(x = cut))
```

```{r}
ggplot(data = ggplot2::diamonds) +
  stat_count(mapping = aes(x = cut))
```

```{r}
demo <- tribble(
  ~a, ~b,
  "bar_1", 20,
  "bar_2", 30,
  "bar_3", 40
)
```


```{r}
ggplot(data = demo) +
  geom_bar(
    mapping = aes(x = a, y = b), stat = "identity"
  )
```

```{r}
ggplot(data = ggplot2::diamonds) +
  geom_bar(
    mapping = aes(x = cut, y = ..prop.., group = 1)
  )
```

```{r}
ggplot(data = ggplot2::diamonds) +
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
  )
```

### Exercises

#### 1

With `stat_summary`:

```{r}
ggplot(data = ggplot2::diamonds) +
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
  )
```

With `geom_pointrange` and a separate stats dataframe:

```{r}
diamonds_summary <- ggplot2::diamonds %>%
  group_by(cut) %>%
  summarise(
    lower = min(depth),
    upper = max(depth),
    p = median(depth)
  )
```

```{r}
ggplot(data = diamonds_summary) +
  geom_pointrange(
    mapping = aes(x = cut, y = p, ymin = lower, ymax = upper)
  )
```

With `geom_pointrange` and in-line stats:

```{r}
ggplot(data = ggplot2::diamonds) +
  geom_pointrange(
    mapping = aes(x = cut, y = depth),
    stat = "summary",
    fun.y = median,
    fun.ymin = min,
    fun.ymax = max
  )
```

With y-axis starting a 0:

```{r}
ggplot(data = ggplot2::diamonds) +
  geom_pointrange(
    mapping = aes(x = cut, y = depth),
    stat = "summary",
    fun.y = median,
    fun.ymin = min,
    fun.ymax = max
  ) +
  scale_y_continuous(limits = c(0, NA))
```

#### 5

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, y = ..prop..))
```

```{r}
ggplot(data = diamonds) +
  geom_bar(
    mapping = aes(x = cut, fill = color, y = ..prop..)
  )
```

```{r}
ggplot(data = ggplot2::diamonds) +
  geom_bar(
    mapping = aes(
      x = cut,
      y = ..prop..,
      fill = color,
      group = 1
    )
  )
```

## Position Arguments

### Total Diamonds by Cut

```{r}
ggplot(data = ggplot2::diamonds) +
  geom_bar(mapping = aes(x = cut, color = cut))
```

```{r}
ggplot(data = ggplot2::diamonds) +
  geom_bar(mapping = aes(x = cut, fill = cut))
```


### Total Diamonds By Cut, By Clarity

```{r}
ggplot(data = ggplot2::diamonds) +
  geom_bar(mapping = aes(x = cut, fill = clarity))
```

```{r}
ggplot(data = ggplot2::diamonds) +
  geom_bar(
    mapping = aes(x = cut, fill = clarity),
    position = "stack"
  )
```


With `position = identity`:

```{r}
ggplot(data = ggplot2::diamonds) +
  geom_bar(
    mapping = aes(x = cut, fill = clarity),
    position = "identity")
```

With `position = identity` and alpha:

```{r}
ggplot(data = ggplot2::diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "identity", alpha = 1/5)
```

With `position = identity` and `fill = NA`:

```{r}
ggplot(data = ggplot2::diamonds) +
  geom_bar(
    mapping = aes(x = cut, color = clarity),
    position = "identity",
    fill = NA
  )
```

With `position = fill`:

```{r}
ggplot(data = ggplot2::diamonds) +
  geom_bar(
    mapping = aes(x = cut, fill = clarity),
    position = "fill"
  )
```

With `position = dodge`:

```{r}
ggplot(data = ggplot2::diamonds) +
  geom_bar(
    mapping = aes(x = cut, fill = clarity),
    position = "dodge"
  )
```

### Highway Mileage (per Gallon) by Engine Displacement (in Litres)

With overplotting:

```{r}
ggplot(data = ggplot2::mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy)
  )
```


With `position = jitter`

```{r}
ggplot(data = ggplot2::mpg) +
  geom_point(
    mapping = aes(x = displ, y = hwy),
    position = "jitter"
  )
```

```{r}
ggplot(data = ggplot2::mpg) +
  geom_jitter(
    mapping = aes(x = displ, y = hwy)
  )
```

### Exercises

#### 1

```{r}
ggplot(data = ggplot2::mpg) +
  geom_point(
    mapping = aes(x = cty, y = hwy),
    position = "jitter"
  )
```

#### 2

```{r}
ggplot(data = ggplot2::mpg) +
  geom_jitter(
    mapping = aes(x = cty, y = hwy),
    width = 0.3, 
    height = 0.3
  )
```

#### 3

```{r}
ggplot(data = ggplot2::mpg) +
  geom_count(
    mapping = aes(x = cty, y = hwy)
  )
```

#### 4

```{r}
ggplot(data = ggplot2::mpg) +
  geom_boxplot(mapping = aes(x = class, y = hwy, color = drv), position = "identity")
```

```{r}
ggplot(data = ggplot2::mpg) +
  geom_boxplot(mapping = aes(x = class, y = hwy, color = drv), position = "dodge")
```


## Coordinate Systems

### coord_flip

```{r}
ggplot(data = ggplot2::mpg) +
  geom_boxplot(mapping = aes(x = class, y = hwy)) +
  coord_flip()
```

### coord_polar

```{r}
bar <- ggplot(data = ggplot2::diamonds) +
  geom_bar(
    mapping = aes(x = cut, fill = cut),
    show.legend = FALSE,
    width = 1
  ) +
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL)
```

```{r}
bar + coord_flip()
```

```{r}
bar + coord_polar()
```

### Exercises

#### 1

```{r}
ggplot(data = ggplot2::diamonds) +
  geom_bar(
    mapping = aes(x = cut, fill = clarity),
    position = "stack"
  )
```

```{r}
ggplot(data = ggplot2::diamonds) +
  geom_bar(
    mapping = aes(x = cut, fill = clarity),
    position = "stack"
  ) + coord_polar()
```

#### 4

```{r}
ggplot(data = ggplot2::mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() +
  coord_fixed() + 
  geom_abline()
```

